---
title: "Organic ML Pretrained Convnet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries

```{r}
library(keras)
```

## Define where the image datasets are

There is a small dataset of 410 images. They are split like this:

|dataset|benzene images|non-benzene images
|---|---|---|
|train|125|125|
|validation|25|25|
|test|55|55|

```{r}
train_dir <- "data/train"
validation_dir <- "data/validation"
test_dir <- "data/test"
```

## Create the image generators

Use data augmentation on the training set, but not the validation set.

```{r}
validation_datagen <- image_data_generator(rescale = 1/255)

train_datagen <- image_data_generator(
  rescale = 1/255,
  rotation_range = 40,
  width_shift_range = 0.2,
  height_shift_range = 0.2,
  shear_range = 0.2,
  zoom_range = 0.2,
  horizontal_flip = TRUE
)

train_generator <- flow_images_from_directory(
  train_dir,
  train_datagen,
  target_size = c(256, 256),
  batch_size = 20,
  class_mode = "binary"
)

validation_generator <- flow_images_from_directory(
  validation_dir,
  validation_datagen,
  target_size = c(256, 256),
  batch_size = 20,
  class_mode = "binary"
)
```

## Define the network

Use VGG16 trained on imagenet for the convolutional layers. Freeze the convolutional layers so that their weights are not adjusted during training. The final layers will be a simple binary classifier made with dense layers.

```{r}
conv_base <- application_vgg16(
  weights = "imagenet",
  include_top = FALSE,
  input_shape = c(256, 256, 3)  # 256x256 1 color channel because images are grayscale
)

conv_base %>% freeze_weights()

network <- keras_model_sequential() %>%
  conv_base %>%
  layer_flatten() %>%
  layer_dense(units = 256, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

network %>% compile(
  loss = "binary_crossentropy",
  optimizer = optimizer_rmsprop(learning_rate = 2e-5),
  metrics = c("accuracy")
)

summary(network)
```

## Fit the network using the training and validation data

```{r}
history <- network %>%
  fit(
    train_generator,
    steps_per_epoch = 100,
    epochs = 30,
    validation_data = validation_generator,
    validation_steps = 50
  )
```

## Save the model

## Save the model

```{r}
network %>% save_model_hdf5("convnet_and_classifier.h5")
```
